# Overview

Note this is a rapid preliminary analysis, so results should not be interpreted as conclusive, as better methods exist to analyse the details.

## Load data

Data are loaded from the repository, while the metadata are directly fetched from the 3D'omics Airtable database.

```{r load_data, message=F, warning=F}
gene_annotations <- read_tsv("data/dram.annotations.tsv.gz") %>% 
  dplyr::rename(gene=1,genome=2,contig=3) %>% 
  mutate(gene = gsub("BGU:[^ ]*BGU:", "BGU:", gene)) %>%  # fix naming error
  mutate(length=end_position-start_position)

gene_counts <- read_tsv("data/gene_counts.tsv") %>% 
  select(-Chr,-Start,-End,-Strand,-Length) %>% 
  pivot_longer(!Geneid, names_to="sample", values_to="raw") %>%
  separate(sample, into=c("sample", "library"),sep="\\.") %>% 
  group_by(Geneid,sample) %>% 
  summarise(raw=sum(raw)) %>% 
  dplyr::rename(gene=Geneid)

# Normalise by gene length
gene_counts <- gene_counts %>% 
  left_join(gene_annotations %>% select(gene,length),by="gene") %>% 
  mutate(norm=raw*150/length) %>% 
  select(gene,sample,raw,norm)

# Transform to relative abundances (per sample)
gene_counts <- gene_counts %>% 
  group_by(sample) %>% 
  mutate(total=sum(norm)) %>% 
  rowwise() %>% 
  mutate(rel=norm/total) %>% 
  select(gene,sample,raw,norm,rel)

genomes <- gene_annotations$genome %>% unique()

sample_metadata <- airtable("tbloNr4f94CbNSka2", "app94YGZqsqwARtWo") %>%
  read_airtable(., fields = c("ID","Bacteria"), id_to_col = TRUE) %>% 
  dplyr::rename(sample=ID) %>% 
  unnest(Bacteria) %>% 
  select(-airtable_record_id)

sample_metadata %>% write_tsv("data/sample_metadata.tsv")

kegg_functions <- read_csv("data/kegg_functions_pivoted_clean.csv")
```

## Basic statistics

### Sequencing depth per sample

```{r overall_stats, message=F, warning=F}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  ggplot(aes(x=raw, y=Bacteria)) +
  geom_col()
```

```{r overall_stats_genes, message=F, warning=F, eval=FALSE}
gene_counts %>% 
  separate(gene,into=c("bin","genome1","genome2"),sep="_",  remove = FALSE) %>% 
  mutate(genome=str_c(genome1," ",genome2)) %>%
  mutate(genome = sub("^(\\w)", "\\U\\1", genome, perl = TRUE)) %>% 
  mutate(genome = str_remove(genome, "@contig")) %>% 
  select(gene,genome,sample,raw) %>% 
  left_join(sample_metadata, by="sample") %>% 
  filter(genome == Bacteria) %>% 
  group_by(Bacteria) %>% 
  summarise(
        genes = n(),                        
        expressed = sum(raw > 1000) 
    ) %>% 
  mutate(percentage=expressed/genes*100) %>% 
  tt()
```

## Carbohydrate active enzymes

### CAZY vs. non-CAZY gene expression

#### Raw reads

```{r cazy_expression_raw, warning=FALSE, comments="", message=FALSE, fig.height=14, fig.width=10, fig.fullwidth=TRUE}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  separate_rows(cazy_ids, sep = "; ")  %>% 
  mutate(cazy=ifelse(!is.na(cazy_ids),"cazy","no_cazy")) %>% 
  select(Bacteria,raw,cazy) %>% 
  group_by(Bacteria,cazy) %>% 
  summarise(raw=sum(raw)) %>% 
  ggplot(aes(x=raw, y=Bacteria, fill=cazy)) +
      geom_col()
```

#### Gene-size normalised

```{r cazy_expression_norm, warning=FALSE, comments="", message=FALSE, fig.height=14, fig.width=10, fig.fullwidth=TRUE}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  separate_rows(cazy_ids, sep = "; ")  %>% 
  mutate(cazy=ifelse(!is.na(cazy_ids),"cazy","no_cazy")) %>% 
  select(Bacteria,norm,cazy) %>% 
  group_by(Bacteria,cazy) %>% 
  summarise(norm=sum(norm)) %>% 
  ggplot(aes(x=norm, y=Bacteria, fill=cazy)) +
      geom_col()
```

### CAZY families

```{r cazy_analysis, warning=FALSE, comments="", message=FALSE, fig.height=18, fig.width=10, fig.fullwidth=TRUE}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  separate_rows(cazy_ids, sep = "; ")  %>% 
  filter(!is.na(cazy_ids)) %>%
  select(Bacteria,cazy_ids,norm) %>% 
  separate(cazy_ids,into="cazy2",sep="_", remove=FALSE) %>%
  group_by(Bacteria,cazy2,norm) %>% 
  summarise(norm=sum(norm)) %>% 
  filter(norm>10) %>% 
  mutate(cazy1=str_sub(cazy2, 1, 2)) %>%
  ggplot(aes(x=Bacteria,y=cazy2,fill=log(norm)))+
    geom_tile() +
    scale_fill_gradient2(low = "white", high ="black") +
    facet_nested(cazy1 ~ ., space="free", scales="free") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## KEGG

```{r top_kegg_overview, message=F, warning=F}
top_genes <- gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  #filter(!is.na(ko_id)) %>% 
  select(Bacteria,ko_id,rel) %>% 
  #filter(Bacteria=="Bacteroides uniformis") %>% 
  group_by(Bacteria,ko_id) %>% 
  summarise(rel=sum(rel)) %>% 
  filter(!is.na(ko_id)) %>% 
  filter(rel>0.001) %>% 
  arrange(-rel) 

top_genes_sort <- top_genes %>% 
  group_by(ko_id) %>% 
  summarise(rel=mean(rel)) %>% 
  arrange(-rel) %>% 
  pull(ko_id)

top_genes %>% 
  mutate(ko_id=factor(ko_id,levels=top_genes_sort)) %>% 
  ggplot(aes(y=Bacteria,x=ko_id,fill=-log(rel))) + 
    geom_tile() + 
    scale_fill_gradient2(low = "white", high ="black")
```

```{r kegg_histogram, message=F, warning=F}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  select(Bacteria,ko_id,rel) %>% 
  filter(Bacteria!="Ruminococcus gnavus") %>% 
  filter(Bacteria!="Mix cultures") %>% 
  group_by(Bacteria,ko_id) %>% 
  summarise(rel=sum(rel)) %>% 
  left_join(kegg_functions, by=join_by("ko_id"=="kegg")) %>% 
  filter(!is.na(ko_id)) %>% 
  filter(rel>0.001) %>% 
  ggplot(aes(y=rel,x=fct_reorder(ko_id,-log(rel)), fill=kegg_function)) + 
    geom_col(width=2) +
    facet_wrap(Bacteria ~ ., scale="free")+
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="bottom") +
  labs(y="Relative abundance",x="KEGG module",fill="Metabolic function") 
```

```{r kegg_distillr_overview, message=F, warning=F}
function_counts <- gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  left_join(kegg_functions, by=join_by("ko_id"=="kegg")) %>% 
  filter(!is.na(kegg_function)) %>% 
  group_by(Bacteria,kegg_function) %>% 
  summarise(norm=sum(norm))

function_counts %>% 
  ggplot(aes(x=norm,y=Bacteria, fill=kegg_function, group=kegg_function)) + 
    geom_bar(stat="identity", colour="white", linewidth=0.1)

```

## 4-bacteria mix

```{r mix4_composition, message=F, warning=F}
gene_counts %>% 
  separate(gene,into=c("bin","genome1","genome2"),sep="_",  remove = FALSE) %>% 
  mutate(genome=str_c(genome1," ",genome2)) %>%
  mutate(genome = sub("^(\\w)", "\\U\\1", genome, perl = TRUE)) %>% 
  left_join(sample_metadata, by="sample") %>% 
  filter(Bacteria=="Mix cultures") %>% 
  group_by(genome) %>% 
  summarise(norm=sum(norm)) %>% 
  filter(norm>1000) %>% 
  ggplot(aes(y=fct_reorder(genome, norm),x=norm)) +
    geom_col()

```
