# Gene selection

## Expression-based

```{r gene_selection_expression, message=F, warning=F}
ko_counts <- gene_counts_relative %>% 
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  left_join(gene_annotations,by="gene") %>%
  group_by(ko_id) %>% 
  summarise(count=mean(count)) %>% 
  arrange(-count)
```

### Around 1%

```{r gene_selection_expression_1, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.01,
         count>0.005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length)) %>% 
  arrange(-length) %>%  
  slice_head(n = 20) %>%
  tt()
```

### Around 0.1%

```{r gene_selection_expression_2, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.001,
         count>0.0005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

### Around 0.01%

```{r gene_selection_expression_3, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.0001,
         count>0.00005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

### Around 0.001%

```{r gene_selection_expression_4, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.00001,
         count>0.000005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

### Around 0.0001%

```{r gene_selection_expression_5, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.000001,
         count>0.0000005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

## Sample-type based

### Flagellum-related genes (K02406, K06603, K02422, K02398, K02407)
- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation

```{r gene_selection_flagellum, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02406|K06603|K02422|K02398|K02407")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Chemotaxis (K03406)

- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation

```{r gene_selection_chemotaxis, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K03406")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```


### Glycoside Hydrolase Family 3 (GH3:K05349)

- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation
- Overexpressed by Prevotella ruminicola in isolation compared to mixture

```{r gene_selection_gh3, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GH3")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Glycoside Hydrolase Family 77 (GH77:K00705)

- Overexpressed by Prevotella ruminicola in isolation compared to mixture

```{r gene_selection_gh77, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GH77")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Gluconate metabolism (K00033, K25031,K07979, K06155)

- Overexpressed by Butyrivibrio fibrosolvens in isolation compared to mixture
- Overexpressed by Megasphaera elsdenii in mixture compared to isolation

```{r gene_selection_gluconate, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K00033|K25031|K07979|K06155")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### ABC transporter (K02003)

- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation

```{r gene_selection_abc, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02003")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Manganese-transport (K12950)
- Overexpressed by Megasphaera elsdenii in isolation compared to mixture 

```{r gene_selection_manganese, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K12950")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Starch-binding (K21572, K21573)
- Overexpressed by Prevotella ruminicola in mixture compared to isolation

```{r gene_selection_starch, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K21572|K21573")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Sortase (K08600)
- Overexpressed by Eubacterium hallii in isolation compared to mixture 

```{r gene_selection_sortase, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K08600")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Thioredoxin (K03671)
- Overexpressed by Eubacterium hallii in isolation compared to mixture 

```{r gene_selection_thioredoxin, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K03671")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

## Biology-based

### Carbohydrate esterase 4 (CE4)

```{r gene_selection_ce4, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "CE4")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Carbohydrate-Binding Module Family 48

```{r gene_selection_cbm48, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "CBM48")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### GlycosylTransferase Family 5

```{r gene_selection_gt5, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GT5")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```

### GlycosylTransferase Family 2

```{r gene_selection_gt2, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GT2")) %>% 
  left_join(gene_counts_relative,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(count>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(count)) %>% 
  arrange(-genes) %>% 
  tt()
```