# Gene selection

## Expression-based

```{r gene_selection_expression, message=F, warning=F}
ko_counts <- gene_counts %>% 
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  left_join(gene_annotations,by="gene") %>%
  group_by(ko_id) %>% 
  summarise(count=mean(rel),
            genes=length(rel),
            prevalence=n_distinct(sample)-1) %>% 
  arrange(-count)
```

### Around 1%

```{r gene_selection_expression_1_long, message=F, warning=F}
ko_counts  %>% 
  filter(count>0.005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(-length) %>%  
  slice_head(n = 20) %>%
  tt()
```

```{r gene_selection_expression_1_short, message=F, warning=F}
ko_counts  %>% 
  filter(count>0.005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(length) %>%  
  slice_head(n = 20) %>%
  tt()
```

### Around 0.1%

```{r gene_selection_expression_2_long, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.001,
         count>0.0005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

```{r gene_selection_expression_2_short, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.001,
         count>0.0005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(length) %>%
  slice_head(n = 30) %>%
  tt()
```

### Around 0.01%

```{r gene_selection_expression_3_long, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.0001,
         count>0.00005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(-length) %>%
  slice_head(n = 50) %>%
  tt()
```

```{r gene_selection_expression_3_short, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.0001,
         count>0.00005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(length) %>%
  slice_head(n = 140) %>%
  tt()
```

### Around 0.001%

```{r gene_selection_expression_4_long, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.00001,
         count>0.000005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

```{r gene_selection_expression_4_short, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.00001,
         count>0.000005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(length) %>%
  slice_head(n = 40) %>%
  tt()
```

### Around 0.0001%

```{r gene_selection_expression_5_long, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.000001,
         count>0.0000005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(count),prevalence=mean(prevalence)) %>% 
  arrange(-length) %>%
  slice_head(n = 20) %>%
  tt()
```

```{r gene_selection_expression_5_short, message=F, warning=F}
ko_counts  %>% 
  filter(count<0.000001,
         count>0.0000005) %>% 
  left_join(gene_annotations,by="ko_id") %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count),length=sum(gene_length),genes=length(ko_id)) %>% 
  arrange(length) %>%
  slice_head(n = 20) %>%
  tt()
```

## Sample-type based

### Flagellum-related genes (K02406, K06603, K02422, K02398, K02407)
- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation

```{r gene_selection_flagellum, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02406|K06603|K02422|K02398|K02407")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_flagellum_K02406, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02406")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_flagellum_K06603, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K06603")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_flagellum_K02398, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02398")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Chemotaxis (K03406)

- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation

```{r gene_selection_chemotaxis, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K03406")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```


### Glycoside Hydrolase Family 3 (GH3:K05349)

- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation
- Overexpressed by Prevotella ruminicola in isolation compared to mixture

```{r gene_selection_gh3, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GH3")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Glycoside Hydrolase Family 77 (GH77:K00705)

- Overexpressed by Prevotella ruminicola in isolation compared to mixture

```{r gene_selection_gh77, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GH77")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Gluconate metabolism (K00033, K25031,K07979, K06155)

- Overexpressed by Butyrivibrio fibrosolvens in isolation compared to mixture
- Overexpressed by Megasphaera elsdenii in mixture compared to isolation

```{r gene_selection_gluconate, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K00033|K25031|K07979|K06155")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gluconate_K00033, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K00033")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gluconate_K25031, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K25031")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```


```{r gene_selection_gluconate_K07979, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K07979")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```


```{r gene_selection_gluconate_K06155, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K06155")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### ABC transporter (K02003)

- Overexpressed by Butyrivibrio fibrosolvens in mixture compared to isolation

```{r gene_selection_abc, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02003")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Manganese-transport (K12950)
- Overexpressed by Megasphaera elsdenii in isolation compared to mixture 

```{r gene_selection_manganese, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K12950")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Starch-binding (K21572, K21573)
- Overexpressed by Prevotella ruminicola in mixture compared to isolation

```{r gene_selection_starch, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K21572|K21573")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Sortase (K08600)
- Overexpressed by Eubacterium hallii in isolation compared to mixture 
```{r gene_selection_sortase, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K08600")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Thioredoxin (K03671)
- Overexpressed by Eubacterium hallii in isolation compared to mixture 

```{r gene_selection_thioredoxin, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K03671")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

## Biology-based

### Carbohydrate esterase 4 (CE4)

```{r gene_selection_ce4, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "CE4")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_ce4_K22278, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K22278")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Carbohydrate-Binding Module Family 48

```{r gene_selection_cbm48, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "CBM48")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### GlycosylTransferase Family 5

```{r gene_selection_gt5, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GT5")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt5_K05366, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K05366")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt5_K00703, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K00703")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt5_K18770, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K18770")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt5_K03814, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K03814")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### GlycosylTransferase Family 2

```{r gene_selection_gt2b, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GT2")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt2_K16698, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K16698")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt2_K02563, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K02563")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt2_K05946, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K05946")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt2_K03429, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K03429")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt2_K00721, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K00721")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gt2_K12999, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K12999")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Glycoside Hydrolase Family 3 (GH3)

```{r gene_selection_gh3b, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GH3")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh3_K01207, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K01207")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh3_K01187, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K01187")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh3_K05349, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K05349")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh3_K12308, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K01811")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh3_K01193, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K01193")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh3_K01191, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K01191")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

### Glycoside Hydrolase Family 77

- Overexpressed by Prevotella ruminicola in isolation compared to mixture

```{r gene_selection_gh77b, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(cazy_ids, "GH77")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh77_K00705, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K00705")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```

```{r gene_selection_gh77_K01208, message=F, warning=F}
gene_annotations %>% 
  filter(str_detect(genome, "butyrivibrio_fibrisolvens|megasphaera_elsdenii|prevotella_ruminicola|anaerobutyricum_hallii")) %>% 
  filter(str_detect(ko_id, "K01208")) %>% 
  left_join(gene_counts,by="gene") %>%
  left_join(sample_metadata,by="sample") %>%
  filter(sample %in% c("C300216","C300205","C300212","C300214","C300203")) %>%
  filter(rel>0) %>% 
  mutate(gene_length=end_position-start_position) %>% 
  group_by(sample,Bacteria) %>% 
  summarise(genes=n(),length_total=sum(gene_length),length_mean=mean(gene_length),count_relative=sum(rel)) %>% 
  arrange(-genes) %>% 
  tt()
```