# BGU strains transcriptomics analysis

Note this is a rapid preliminary analysis, so results should not be interpreted as conclusive, as better methods exist to analyse the details.

```{r load_libraries, message=F, warning=F}
library(tidyverse)
library(ggplot2)
library(rairtable)
library(OmicCircos)
library(tinytable)
library(ggh4x)
```

## Load data

Data are loaded from the repository, while the metadata are directly fetched from the 3D'omics Airtable database.

```{r load_data, message=F, warning=F}
gene_counts <- read_tsv("data/subread.REF0018-bgu_genomes-drep.1.00.tsv.gz") %>% 
  pivot_longer(!sequence_id, names_to="sample", values_to="count") %>%
  separate(sample, into=c("sample", "library"),sep="\\.") %>% 
  group_by(sequence_id,sample) %>% 
  summarise(count=sum(count)) %>% 
  dplyr::rename(gene=sequence_id)

gene_counts_relative <- read_tsv("data/subread.REF0018-bgu_genomes-drep.1.00.tsv.gz") %>%
  mutate_at(vars(-sequence_id),~./sum(.)) %>%
  pivot_longer(!sequence_id, names_to="sample", values_to="count") %>%
  separate(sample, into=c("sample", "library"),sep="\\.") %>% 
  group_by(sequence_id,sample) %>% 
  summarise(count=sum(count)) %>% 
  dplyr::rename(gene=sequence_id)

gene_annotations <- read_tsv("data/dram.annotations.tsv.gz") %>% 
  dplyr::rename(gene=1,genome=2,contig=3)

genomes <- gene_annotations$genome %>% unique()

sample_metadata <- airtable("tbloNr4f94CbNSka2", "app94YGZqsqwARtWo") %>%
  read_airtable(., fields = c("ID","Bacteria"), id_to_col = TRUE) %>% 
  dplyr::rename(sample=ID) %>% 
  unnest(Bacteria) %>% 
  select(-airtable_record_id)

sample_metadata %>% write_tsv("data/sample_metadata.tsv")

kegg_functions <- read_csv("data/kegg_functions_pivoted_clean.csv")
```

## Basic statistics

```{r overall_stats, message=F, warning=F}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  ggplot(aes(x=count, y=Bacteria)) +
  geom_col()
```
```{r overall_stats_genes, message=F, warning=F}
gene_counts %>% 
  separate(gene,into=c("bin","genome1","genome2"),sep="_",  remove = FALSE) %>% 
  mutate(genome=str_c(genome1," ",genome2)) %>%
  mutate(genome = sub("^(\\w)", "\\U\\1", genome, perl = TRUE)) %>% 
  select(gene,genome,sample,count) %>% 
  left_join(sample_metadata, by="sample") %>% 
  filter(genome == Bacteria) %>% 
  group_by(Bacteria) %>% 
  summarise(
        genes = n(),                        
        expressed = sum(count > 1000) 
    ) %>% 
  mutate(percentage=expressed/genes*100) %>% 
  tt()
```

## Carbohydrate active enzymes

### CAZY vs. non-CAZY gene expression

```{r cazy_expression, warning=FALSE, comments="", message=FALSE, fig.height=14, fig.width=10, fig.fullwidth=TRUE}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  separate_rows(cazy_ids, sep = "; ")  %>% 
  mutate(cazy=ifelse(!is.na(cazy_ids),"cazy","no_cazy")) %>% 
  select(Bacteria,count,cazy) %>% 
  group_by(Bacteria,cazy) %>% 
  summarise(count=sum(count)) %>% 
  ggplot(aes(x=count, y=Bacteria, fill=cazy)) +
      geom_col()
```

### CAZY families

```{r cazy_analysis, warning=FALSE, comments="", message=FALSE, fig.height=14, fig.width=10, fig.fullwidth=TRUE}
gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  separate_rows(cazy_ids, sep = "; ")  %>% 
  filter(!is.na(cazy_ids)) %>%
  select(Bacteria,cazy_ids,count) %>% 
  separate(cazy_ids,into="cazy2",sep="_", remove=FALSE) %>%
  group_by(Bacteria,cazy2,count) %>% 
  summarise(count=sum(count)) %>% 
  filter(count>1000) %>% 
  mutate(cazy1=str_sub(cazy2, 1, 2)) %>%
  ggplot(aes(x=Bacteria,y=cazy2,fill=log(count)))+
    geom_tile() +
    scale_fill_gradient2(low = "white", high ="black") +
    facet_nested(cazy1 ~ ., space="free", scales="free") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## KEGG

```{r top_kegg_overview, message=F, warning=F}
top_genes <- gene_counts_relative %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  #filter(!is.na(ko_id)) %>% 
  select(Bacteria,ko_id,count) %>% 
  #filter(Bacteria=="Bacteroides uniformis") %>% 
  group_by(Bacteria,ko_id) %>% 
  summarise(count=sum(count)) %>% 
  filter(!is.na(ko_id)) %>% 
  filter(count>0.001) %>% 
  arrange(-count) 

top_genes_sort <- top_genes %>% 
  group_by(ko_id) %>% 
  summarise(count=mean(count)) %>% 
  arrange(-count) %>% 
  pull(ko_id)

top_genes %>% 
  mutate(ko_id=factor(ko_id,levels=top_genes_sort)) %>% 
  ggplot(aes(y=Bacteria,x=ko_id,fill=-log(count))) + 
    geom_tile() + 
    scale_fill_gradient2(low = "white", high ="black")
```

```{r kegg_histogram, message=F, warning=F}
gene_counts_relative %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  select(Bacteria,ko_id,count) %>% 
  filter(Bacteria!="Ruminococcus gnavus") %>% 
  filter(Bacteria!="Mix cultures") %>% 
  group_by(Bacteria,ko_id) %>% 
  summarise(count=sum(count)) %>% 
  left_join(kegg_functions, by=join_by("ko_id"=="kegg")) %>% 
  filter(!is.na(ko_id)) %>% 
  filter(count>0.001) %>% 
  ggplot(aes(y=count,x=fct_reorder(ko_id,-log(count)), fill=kegg_function)) + 
    geom_col(width=2) +
    facet_wrap(Bacteria ~ ., scale="free")+
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="bottom") +
  labs(y="Relative abundance",x="KEGG module",fill="Metabolic function") 
```

```{r kegg_distillr_overview, message=F, warning=F}
function_counts <- gene_counts %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  left_join(kegg_functions, by=join_by("ko_id"=="kegg")) %>% 
  filter(!is.na(kegg_function)) %>% 
  group_by(Bacteria,kegg_function) %>% 
  summarise(count=sum(count))

function_counts %>% 
  ggplot(aes(x=count,y=Bacteria, fill=kegg_function, group=kegg_function)) + 
    geom_bar(stat="identity", colour="white", linewidth=0.1)

```

## 4-bacteria mix

```{r mix4_composition, message=F, warning=F}
gene_counts %>% 
  separate(gene,into=c("bin","genome1","genome2"),sep="_",  remove = FALSE) %>% 
  mutate(genome=str_c(genome1," ",genome2)) %>%
  mutate(genome = sub("^(\\w)", "\\U\\1", genome, perl = TRUE)) %>% 
  left_join(sample_metadata, by="sample") %>% 
  filter(Bacteria=="Mix cultures") %>% 
  group_by(genome) %>% 
  summarise(count=sum(count)) %>% 
  filter(count>10000) %>% 
  ggplot(aes(y=fct_reorder(genome, count),x=count)) +
    geom_col()

```

### Mix vs. isolates

```{r mix4_relabun, message=F, warning=F}
isolates_relabun <- gene_counts %>% 
  separate(gene,into=c("bin","genome1","genome2"),sep="_",  remove = FALSE) %>% 
  mutate(genome=str_c(genome1," ",genome2)) %>%
  mutate(genome = sub("^(\\w)", "\\U\\1", genome, perl = TRUE)) %>% 
  select(gene,genome,sample,count) %>% 
  left_join(sample_metadata, by="sample") %>% 
  filter(genome == Bacteria) %>% 
  group_by(genome) %>% 
  group_split() %>%
  lapply(function(df) {
    df %>% mutate(count = count / sum(count, na.rm = TRUE))
  }) %>%
  setNames(map_chr(., ~ unique(.x$genome)))

mix_relabun <- gene_counts %>% 
  separate(gene,into=c("bin","genome1","genome2"),sep="_",  remove = FALSE) %>% 
  mutate(genome=str_c(genome1," ",genome2)) %>%
  mutate(genome = sub("^(\\w)", "\\U\\1", genome, perl = TRUE)) %>% 
  left_join(sample_metadata, by="sample") %>% 
  filter(Bacteria=="Mix cultures") %>% 
  filter(genome %in% c("Butyvibrio fibrisolvens","Megasphaera elsdenii","Prevotella ruminicola","Eubacterium hallii")) %>%
  select(gene,genome,Bacteria,sample,count) %>% 
  group_by(genome) %>% 
  group_split() %>%
  lapply(function(df) {
    df %>% mutate(count = count / sum(count, na.rm = TRUE))
  }) %>%
  setNames(map_chr(., ~ unique(.x$genome)))
```

#### Megasphaera elsdenii

```{r isolate_mix_ME_dif}
isolates_relabun$`Megasphaera elsdenii` %>% 
  select(gene,count) %>% 
  left_join(mix_relabun$`Megasphaera elsdenii`,by="gene") %>% 
  dplyr::rename(isolate=count.x, mix=count.y) %>% 
  mutate(difference=isolate-mix) %>% 
  arrange(difference) %>%
  { bind_rows(head(., 20), tail(., 20)) } %>%
  left_join(gene_annotations, by="gene") %>% 
  ggplot(aes(x=difference,y=ko_id)) +
    geom_point()
  
isolates_relabun$`Megasphaera elsdenii` %>% 
  select(gene,count) %>% 
  left_join(mix_relabun$`Megasphaera elsdenii`,by="gene") %>% 
  dplyr::rename(isolate=count.x, mix=count.y) %>% 
  mutate(difference=isolate-mix) %>% 
  arrange(difference) %>%
  { bind_rows(head(., 20), tail(., 20)) } %>%
  left_join(gene_annotations, by="gene") %>% 
  ggplot(aes(x=difference,y=cazy_ids)) +
    geom_point()
```

```{r isolate_mix_ME_plot, message=F, warning=F}
bind_rows(
  isolates_relabun$`Megasphaera elsdenii` %>% mutate(type="isolate"),
  mix_relabun$`Megasphaera elsdenii` %>% mutate(type="mix")) %>% 
  ggplot(aes(y=gene,x=type,fill=count)) +
    geom_tile()+
    scale_fill_gradient2(low = "white", high ="black") +
    theme_classic() +
    theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
#### Prevotella ruminicola

```{r isolate_mix_PR_dif}
isolates_relabun$`Prevotella ruminicola` %>% 
  select(gene,count) %>% 
  left_join(mix_relabun$`Prevotella ruminicola`,by="gene") %>% 
  dplyr::rename(isolate=count.x, mix=count.y) %>% 
  mutate(difference=isolate-mix) %>% 
  arrange(difference) %>%
  { bind_rows(head(., 20), tail(., 20)) } %>%
  left_join(gene_annotations, by="gene") %>% 
  ggplot(aes(x=difference,y=ko_id)) +
    geom_point()
  
isolates_relabun$`Prevotella ruminicola` %>% 
  select(gene,count) %>% 
  left_join(mix_relabun$`Prevotella ruminicola`,by="gene") %>% 
  dplyr::rename(isolate=count.x, mix=count.y) %>% 
  mutate(difference=isolate-mix) %>% 
  arrange(difference) %>%
  { bind_rows(head(., 20), tail(., 20)) } %>%
  left_join(gene_annotations, by="gene") %>% 
  ggplot(aes(x=difference,y=cazy_ids)) +
    geom_point()
```

```{r isolate_mix_PR_plot, message=F, warning=F}
bind_rows(
  isolates_relabun$`Prevotella ruminicola` %>% mutate(type="isolate"),
  mix_relabun$`Prevotella ruminicola` %>% mutate(type="mix")) %>% 
  ggplot(aes(y=gene,x=type,fill=count)) +
    geom_tile()+
    scale_fill_gradient2(low = "white", high ="black") +
    theme_classic() +
    theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r expression_circos, message=F, warning=F, echo=FALSE, eval=FALSE}
segments <- gene_annotations %>% 
  filter(genome=="BGU:bin_bacteroides_uniformis") %>% 
  group_by(contig) %>% 
  summarise(fragment=max(end_position)) %>% 
  ungroup() %>% 
  mutate(chromStart = lag(cumsum(fragment), default = 0) + 1, chromEnd = cumsum(fragment)) %>% 
  dplyr::rename(chrom=1) %>% 
  select(chrom,chromStart,chromEnd)

segments_angle <- segAnglePo(seg.dat=segments,seg=segments$chrom)

mapping <- gene_counts_relative %>% 
  left_join(sample_metadata, by="sample") %>% 
  left_join(gene_annotations, by="gene") %>% 
  filter(sample=="C300204") %>% 
  select(contig,start_position,count,ko_id,gene) %>% 
  filter(contig %in% segments$chrom) %>% 
  mutate(count=-log(count)) %>% 
  dplyr::rename(chrom=1,po=2)

plot(c(1,800), c(1,800), type="n", axes=FALSE, xlab="", ylab="", main="")
circos(R=260, cir=segments_angle, W=20, mapping=mapping, col.v=3, type="ss", B=TRUE, col=colors[1], lwd=0.1, scale=TRUE)

circos(R=260, cir=segments_angle, W=40, mapping=mapping, col.v=3, type="hist", B=TRUE, scale=TRUE)


  
  group_by(Bacteria,contig) %>% 
  summarise(fragment=max(end_position)) %>% 
  ungroup() %>% 
  mutate(chromStart = lag(cumsum(fragment), default = 0) + 1, chromEnd = cumsum(fragment)) %>% 
  dplyr::rename(chrom=1) %>% 
  select(chrom,chromStart,chromEnd)


```
